---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sasfunclust

<!-- badges: start -->
<!-- badges: end -->

The package  sasfunclust implements the the sparse and smooth functional clustering (SaS-Funclust) method proposed by Centofanti et al. (2021).
SaS-Funclust is a new method for clustering functional data that aims to classify a sample of curves into homogeneous groups while jointly detecting the most informative portions of domain.
The method relies on a general functional Gaussian mixture model whose parameters are estimated by maximizing a log-likelihood function penalized with the functional adaptive pairwise fusion penalty (FAPFP) and a roughness penalty.
The package comprises two main functions \code{sasfclust} and \code{sasfclust_cv}. 
The former performs the SaS-Funclust for fixed number of clusters \eqn{G}, tuning parameter of the smoothness penalty \eqn{\lambda_s}, and tuning parameter of the FAPFP \eqn{\lambda_l}.
The latter executes the K-fold cross-validation procedure  described in Centofanti et al. (2021) to   choose  \eqn{G}, \eqn{\lambda_s}, and \eqn{\lambda_l}.

## Installation



The development version can be installed from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fabiocentofanti/sasfunclust")
```
## Example


This is a basic example which shows you how to apply the two main functions \code{sasfclust} and \code{sasfclust_cv} on a synthetic dataset generated as as described in the simulation study of Centofanti et al. (2021).



We start by loading and attaching the sasfunclust package.
```{r }
library(sasfunclust)
```

Then, we generate the synthetic dataset as follows.
```{r }
n_i=20
train<-simulate_data("Scenario I",n_i=n_i,var_e = 1,var_b = 0.5^2)
```

To apply \code{sasfclust_cv},  sequences of \eqn{G}, \eqn{\lambda_s}, and \eqn{\lambda_l} should be defined.
```{r }
lambda_s_seq=10^seq(-4,-3)
lambda_l_seq=10^seq(0,3)
G_seq=2
```
And, then, \code{sasfclust_cv} is executed.
```{r }
mod_cv<-sasfclust_cv(X=train$X,grid=train$grid,G_seq=G_seq,
lambda_l_seq = lambda_l_seq,lambda_s_seq =lambda_s_seq,maxit = 200,K_fold = 5,q=30,ncores=8)
```

The result are plotted.
```{r, dev='pdf',dpi=1000}
plot(mod_cv)
```
By using the model selection method described in Centofanti et al. (2021), the optimal values of \eqn{G}, \eqn{\lambda_s}, and \eqn{\lambda_l}, are \eqn{2}, \eqn{10^-4}, and \eqn{10^2}, respectively.

Finally, \code{sasfclust} is applied with \eqn{G}, \eqn{\lambda_s}, and \eqn{\lambda_l} fixed to their optimal values.
```{r }
mod<-sasfclust(X=train$X,grid=train$grid,G=mod_cv$G_opt,
lambda_l = mod_cv$lambda_l_opt,lambda_s =mod_cv$lambda_s_opt,maxit = 200,q=30)
```
The cluster membership vector and the plot of the estimated cluster mean functions and the classified curves are obtained as follows.
```{r example}
print(mod$clus$classes)
plot(mod)
```
# References


